\pkg{pksensi}: An R package to apply sensitivity analysis in pharmacokinetic models

### 2019-06-26 manuscript/manuscript.Rmd
Concluding

In our previous work \citep{fphar201800588}, we developed an approach to apply global sensitivity analysis (GSA) workflow to reduce the computational burden in the Bayesian,  Markov Chain Monte Carlo (MCMC)-based calibration process of a physiologically based pharmacokinetic (PBPK) model. We used GNU MCSim \citep{JSSv002i09}, an effective simulation package for Bayesian population PBPK modeling, to calibrate the model. We found that the extended Fourier Amplitude Sensitivity Test (eFAST), a type of variance-based GSA algorithm, had the best balance of efficiency and accuracy for a complex, multi-compartment, multi-dataset, and multi-metabolite PBPK model. It's more comptational efficient in calculate the influence of model parameter than other GSA methods via Monte Carlo integration, such as Jansen or OWen method. Also, we found some efficient visualization approaches that can be used to distinguish between “influential” and “non-influential” parameters. We also found a useful approach for communicating the parameter sensitivity in decision making.

@@

*	Cross-platform: Models can run on Windows/MacOS/Linux
*	Freely available: All related packages are free and open source
*	Integrated application: Users can run pharmacokinetic models in \proglang{R} with script that were written in \proglang{C} or \pkg{GNU MCSim} 
*	Decision making: The output results and visualization tools can be used to easily determine which parameters have "non-influential" effects on the model output and can be fixed in model calibration.

@@

The \pkg{GNU MCSim} is a powerful simulation software package, that can conduct various dynamical simulations for biological-based modeling, and it's mainly built for pharmacokinetic research with probabilistic approaches [@bois2009gnu]. The \Proglang{R} is a comprehensive statistical computing language that can conduct probabilistic programming and have tons of packages for statistical analysis and data visualization.  Some fundamental packages on the CRAN repository and other software platforms hasd been built to perform different types of global sensitivity analysis. The development of \pkg{pksensi} is mainly for physiologically based kinetic models that are constructed by differential equations. It provides a straightforward application to investigate the impact of parameters on model output across time and variables. We adopted the eFSAT method, a common variance-based approach and had better efficiency and accuracy in the model assessment.

in \pkg{pksensi}

  have the developed solver to solve the differential equation in dynamic simulation.

The dosing schedule

Some studies applied another GSA algorithm in the physiologically and biologically based ODE model, such as compartmental absorption and transit (also known as CAT) and other system biology model [@melillo2019variance; @sumner2012methodology]. In future version, we expect to add additional GSA approach that had been applied in advance study.

Principle component

>>>>

In our previous work \citep{fphar201800588}, we developed an approach to apply global sensitivity analysis (GSA) workflow to reduce the computational burden in the Bayesian,  Markov Chain Monte Carlo (MCMC)-based calibration process of a physiologically based pharmacokinetic (PBPK) model. We used GNU MCSim \citep{JSSv002i09}, an effective simulation package for Bayesian population PBPK modeling, to calibrate the model. We found that the extended Fourier Amplitude Sensitivity Test (eFAST), a type of variance-based GSA algorithm, had the best balance of efficiency and accuracy for a complex, multi-compartment, multi-dataset, and multi-metabolite PBPK model. It's more comptational efficient in calculate the influence of model parameter than other GSA methods via Monte Carlo integration, such as the GSA approach from \citep{jansen1999analysis} and \citep{owen2013better}. Also, we found some efficient visualization approaches that can be used to distinguish between “influential” and “non-influential” parameters. We also found a useful approach for communicating the parameter sensitivity in decision making.

@@

*	Cross-platform: Models can run on Windows/MacOS/Linux
*	Freely available: All related packages are free and open source
*	Integrated application: Users can run pharmacokinetic models in \proglang{R} with script that were written in \proglang{C} or \pkg{GNU MCSim} 
*	Decision support: The output results and visualization tools can be used to easily determine which parameters have "non-influential" effects on the model output and can be fixed in model calibration.

@@

The \pkg{GNU MCSim} is a powerful simulation software package, that can conduct various dynamical simulations for biological-based modeling, and it's mainly built for pharmacokinetic research with probabilistic approaches [@bois2009gnu]. The \proglang{R} is a comprehensive statistical computing language that can conduct probabilistic programming and have tons of packages for statistical analysis and data visualization. The current version had included \pkg{GNU MCSim} and \pkg{deSolve} as the main solver in our model assessment. Compare with \pkg{deSolve} package, we found that The \pkg{GNU MCSim} can provide better efficiency to solve the ODE in PK model. Some developed package in \proglang{R} had also provide the feature to solve ODE such as \pkg{RxODE} and \pkg{mrgsolve} [@wang2016tutorial; @R-mrgsolve]. Therefore, we hope to incorporate this package in our GSA workflow to provide user more options in the modeling.

Moreover, some fundamental packages on the CRAN repository and other software platforms had been built to perform different types of GSA. The development of \pkg{pksensi} is mainly for physiologically based kinetic models that are constructed by differential equations. It provides a straightforward application to investigate the impact of parameters on model output across time and variables. We adopted the eFSAT method, a common variance-based sensitivity analysis that had better efficiency and accuracy in the model assessment. It can exam the parameter effect from first order to the total order. In addition, the package can not only conduct GSA but also include the function to assess the convergence of sensitivity measurements, which is rarely discussed in most GSA studies and its software package. We also develop the function to visualize the output result that can help us determine the "influential" and "noninfluential" parameters that can be applied in parameter fixing in the model Calibration. In addition to eFAST, Some studies applied another GSA algorithm in the physiologically and biologically based ODE model, such as compartmental absorption and transit (also known as CAT) and other system biology model [@melillo2019variance; @sumner2012methodology]. In future version, we expect to add additional GSA approach that had been applied in advance study. The other features such as principle component analysis [@lamboni2011multivariate] were considered to include in \pkg{pksensi}.

It is worth noting that the GSA results are limited to the specific model and the given range of the model parameter. The provided conditions, such as the number of examing parameter, parameter distribution, parameter range, and the given dosing level in PK model might have a potential impact on the level of importance of each parameter. The misconduct of the setting condition might lead to the incorrect estimated result. In other words, if we have various types of experiment data (e.g. different exposure routes and biological measurements) and will use them in model calibration, it will enhance the difficulties in GSA. The complex model with a high dimension of the parameter is also a challenge in GSA, which will need higher computer demand in the model assessment. Overall, we have developed \pkg{pksensi} that can apply the comprehensive GSA workflow for most PK models. More additional applications still need to be investigated and strengthen.
